---
title: "R Notebook"
output: html_notebook
---

# ```{r}
# data <- data.frame(unlist(X2016FallSemWireless))
# ```


Function: Removes staff
Purpose: Staff don't collaborate the way other people do
```{r}
print(length(X2016FallSemWireless[,1]))
wirelessData <- X2016FallSemWireless[filteredFallData$campus != 'cuc',]
print(length(wirelessData[,1]))
```


Functionality: Plots density curves for each location's signal strenth.
Purpose: Figure out whether there are certain places where the signal strenth is bad (possibly people connecting to another floor)

```{r}
plotSigStrength <- function(dataset) {
  j = 0;
  myList <- split(dataset, dataset$WAPID)
  for (i in levels(as.factor(dataset$WAPID))) {
    # if (j < 4) {
      my1 <- myList[[i]]
      signal <- na.omit(my1$avgSigQual)
      if (length(signal) > 1) {
        d <- density(na.omit(my1$avgSigQual))
        plot(d, main=i)
      } else {
        print(i)
      } 
    }
    j = j+1
  # }
}

```


Functionality: Density curves by location, all data (except staff)
Purpose: try to figure out if people are connecting to other floors
Results: Most access points have two distince points; one good, one bad. A few access points are just bad
```{r}
plotSigStrength(wirelessData)
```

**Functionality**: Check whether connection bytes and time of 0.0 connections differs  
**Purpose**: We'd expect people with a zero connection to have much fewer bytes per sec  
**Result**: Our predictions are correct, although not as much as we had expected.  
But more importantly, it looks like we've got a bunch of connections with kbps=0.
Unclear what we should do with this. We should figure out whether this represents actual
people in the library who just aren't using their computer or whether these are errors.  
```{r}
nonzeroConnection <- ifelse(wirelessData$avgSigQual > 0.0, "Nonzero Connection Strength", "Connection Strength = 0")
# colnames(nonzeroConnection) = c("Connection Strength = 0", "Nonzero Connection Strength")
print(length(nonzeroConnection))
boxplot(wirelessData$avgKbps ~ nonzeroConnection, main="Kilobytes Per Second by Average Connection Strength", xlab="", ylab="Average Kbps")
print("done")
```

**Functionality**: Trims out people with kbps=0 (just to test)  
**Purpose**: Want to verify the thing we were trying to do above - checking how
connectivity patterns differ based on signal quality  
**Result**: Graphs look almost identical.  This is a good sign! 
It means we have a bunch of outliers, but that it's not like most of our data is just useless.
When we zoom in on that range, we see there are diff connection patterns based on signal
quality.  This makes sense.  
```{r}
print(length(wirelessData[,1]))
wirelessData1a <- wirelessData[wirelessData$avgKbps > 0.0,]
print(length(wirelessData1a[,1]))

nonzeroConnection <- (wirelessData1a$avgSigQual > 0.0)
print(length(nonzeroConnection))
boxplot(wirelessData1a$avgKbps ~ nonzeroConnection)

# Try again, but now cut out big values
print(length(wirelessData[,1]))
wirelessData2a <- wirelessData[wirelessData$avgKbps < 300,]
wirelessData2a <- wirelessData2a[wirelessData2a$avgKbps > 0,]
print(length(wirelessData2a[,1]))

nonzeroConnection <- ifelse(wirelessData2a$avgSigQual > 0.0, "Nonzero Connection Strength", "Connection Strength = 0",
                            main="")
print(length(nonzeroConnection))
boxplot(wirelessData2a$avgKbps ~ nonzeroConnection)

```

**Functionality**: Trims data which contains connection times over 8 hours, which doesn't seem reasonable. 
Then makes boxplots of duration time for people with zero/nonzero connection strength.  
**Purpose**: We want to see if we should exclude the people with bad connections.  Also, it would be useful
to see how many connection times are wonky.  
**Result**: Over 10 percent of our data set had bad connection times! We should fix that.
Also, looks like people w 0 connection strength do connect shorter, but not THAT much shorter.  

```{r}
# Remove data points with unreasonable connection times
print(length(wirelessData[,1]))
wirelessData3a <- wirelessData[wirelessData$duration_mins < 480,]
wirelessData3a <- wirelessData3a[wirelessData3a$duration_mins >= 0,]
print(length(wirelessData3a[,1]))

nonzeroConnection <- (wirelessData3a$avgSigQual > 0.0)
print(length(nonzeroConnection))
boxplot(wirelessData3a$duration_mins ~ nonzeroConnection)
```






**Functionality**: Trim out wireless data with signal = 0  
**Purpose**: This data is unlikely to be an accurate representation of where this person is working. 
We should verify this with the library guy though.  
**Result**: About a quarter of our observations had 0 signal strength  
```{r}
print(length(wirelessData[,1]))
wirelessData1 <- wirelessData[wirelessData$avgSigQual != 0.0,]
print(length(wirelessData1[,1]))
plotSigStrength(wirelessData1)
```

```{r}
i = 0
calculateDuration <- function(r) {
  print(r[20])
  duration = r[20] - r[12]
  duration = duration + (r[19] - r[11]) * 60
  # duration = duration + (r$disconnect_day - r$connect_day) * 24 * 60
  if (r[17] != r[10]) {
    if(r[17] > r[10]) {
      duration = duration + (r[17] - r[10]) * 24 * 60
    } 
      #else {
    #   # Not all months have thirty days, but if we're already on this stage,
    #   # Then something is already highly weird
    #   duration = duration + (as.numeric(r[16]) - as.numeric(r[9])) * 30 * 24 * 60
    # }
  }
  i = i + 1
  return(duration)
}
```


```{r}
wirelessData5a <- wirelessData3a
wirelessData5a$disconnect_day <- ifelse(wirelessData5a$disconnect_hour == '0', wirelessData5a$disconnect_day + 1, wirelessData5a$disconnect_day)
  
wirelessData5a$connect_day <- ifelse(wirelessData5a$connect_hour == '0', wirelessData5a$connect_day + 1, wirelessData5a$connect_day)
```


```{r}
get_doy <- function(x) {
  switch(as.character(x),
         "7" = 182,
         "8" = 213,
         "9" = 244,
         "10" = 274,
         "11" = 305,
         "12" = 335,
         {
           print(x)
           x
         })
}
```


```{r}
wirelessData2b <- na.omit(wirelessData2b)
wirelessData2b$disconnect_doy <- unlist(sapply(wirelessData2b$disconnect_month, get_doy)) + wirelessData2b$disconnect_day
```



```{r}
# Convert 0 to 24
wirelessData2 <- wirelessData

# wirelessData2$disconnect_day <- ifelse(wirelessData2$disconnect_hour == '0', wirelessData2$disconnect_day + 1, wirelessData2$disconnect_day)
# 
# wirelessData2$connect_day <- ifelse(wirelessData2$connect_hour == '0', wirelessData2$connect_day + 1, wirelessData2$connect_day)

# Re-calculate duration
colLen = length(wirelessData[,1])

wirelessData2b <- wirelessData2
wirelessData2b$connect_minute <- as.numeric(wirelessData2b$connect_minute)
wirelessData2b$connect_hour <- as.numeric(wirelessData2b$connect_hour)
wirelessData2b$connect_day <- as.numeric(wirelessData2b$connect_day)
wirelessData2b$connect_month <- as.numeric(wirelessData2b$connect_month)

wirelessData2b$disconnect_minute <- as.numeric(wirelessData2b$disconnect_minute)
wirelessData2b$disconnect_hour <- as.numeric(wirelessData2b$disconnect_hour)
wirelessData2b$disconnect_day <- as.numeric(wirelessData2b$disconnect_day)
wirelessData2b$disconnect_month <- as.numeric(wirelessData2b$disconnect_month)

# wirelessData2b$disconnect_hour <- ifelse(
#   (wirelessData2b$disconnect_hour == 0 || wirelessData2b$disconnect_hour == 12) ,
#   (wirelessData2b$disconnect_hour + 12),
#   wirelessData5a$disconnect_hour)

# wirelessData2b$disconnect_hour <- ifelse(
#   (wirelessData2b$disconnect_hour == 12 || wirelessData2b$disconnect_hour == 0),
#   wirelessData5a$disconnect_hour + 12,
#   wirelessData5a$disconnect_hour)
# 
wirelessData2b$disconnect_hour <- ifelse(
  wirelessData2b$disconnect_hour == 12,
  24,
  wirelessData2b$disconnect_hour)


wirelessData2b$disconnect_hour <- ifelse(
  wirelessData2b$disconnect_hour == 0,
  12,
  wirelessData2b$disconnect_hour)
# 

# wirelessData2b$connect_hour <- ifelse(
#   (wirelessData2b$connect_hour == 0 || wirelessData2b$connect_hour == 12) , 
#   (wirelessData2b$connect_hour + 12), 
#   wirelessData5a$connect_hour)

print(length(wirelessData2b[,1]))
wirelessData2b <- na.omit(wirelessData2b)
print(length(wirelessData2b[,1]))

wirelessData2b$disconnect_doy <- unlist(sapply(wirelessData2b$disconnect_month, get_doy)) + wirelessData2b$disconnect_day

wirelessData2b$duration_mins <- wirelessData2b$disconnect_minute - wirelessData2b$connect_minute +
      (wirelessData2b$disconnect_hour - wirelessData2b$connect_hour) * 60 +
      (wirelessData2b$disconnect_doy - wirelessData2b$connect_doy) * 60 * 24

# A couple weird cases overlapping the start of daylight savings time
# wirelessData2b$duration_mins[307133:307138] <- wirelessData2b$duration_mins[307133:307138] + 60

# Remove cases with connection tim
     

```


```{r}
d <- density(wirelessData2b$duration_mins)
plot(d, main="Density Plot of Connection Durations in Minutes")
hist(wirelessData2b$duration_mins)

wirelessData2c <- wirelessData2b[wirelessData2b$duration_mins < 600,]
d2 <- density(wirelessData2c$duration_mins)
plot(d2, main="Density Plot of Connection Durations; Max 600 min")
# 
# plot(wirelessData2c$duration_mins, wirelessData2c$avgSigQual, main="Insert useful title")
# plot(wirelessData2b$duration_mins, wirelessData2b$avgSigQual, main="Insert useful title")
# 
# 
# plot(wirelessData2c$duration_mins, wirelessData2c$avgKbps, main="Insert useful title")
plot(wirelessData2b$duration_mins, wirelessData2b$avgKbps, main="Kbps vs Connection Duration", xlab="Connection Duration in Minutes", ylab="Average Kbps")

```

Interesting.  A fair number of big ones.  Let's get rid of them.


```{r}
wirelessData0 <- wirelessData2c[wirelessData2c$disconnect_hour == 0,]
wirelessData12 <- wirelessData2c[wirelessData2c$disconnect_hour == 12,]
d0 <- density(wirelessData0$duration_mins)
plot(d0)
d12 <- density(wirelessData12$duration_mins)
plot(d12)

wirelessData11 <- wirelessData2c[wirelessData2c$disconnect_hour == 11,]
d11 <- density(wirelessData11$duration_mins)
plot(d11)

# plot(wirelessData2c$duration_mins, wirelessData2c$avgSigQual, main="Insert useful title")
```

WHAT.  We should... fix this?

```{r}
# A couple weird cases overlapping the start of daylight savings time
wirelessData2b$duration_mins[307133:307138] <- wirelessData2b$duration_mins[307133:307138] + 60


```


```{r}
# Summary of data Analysis
vec = c(689203, 601074, 560763, 543141)
labs = c("Initial", "Staff Removed", "NAs Omitted", "Long Connections Removed")
help <- data.frame()
help$vec = vec
help$labs = labs
plot(vec ~ labs, help)
```





**Functionality**: Plot connection times, see if any seem especially weird.
**Purpose**: Seee if there are any which should be omitted
** Results**: There are a few outliers. Let's cutoff at 480 (10 hrs). Also, what???

```{r}
wirelessData4a <- wirelessData[wirelessData$duration_mins >= 0,]
d <- density(wirelessData4a$duration_mins)
plot(d)
d2 <- density(wirelessData3a$duration_mins)
plot(d2)
```

```{r}
plot(wirelessData2a$avgKbps, wirelessData2a$duration_mins, main="Scatterplot Example")

plot(wirelessData3a$duration_mins, wirelessData3a$avgKbps, main="Scatterplot Example")

plot(wirelessData3a$duration_mins, wirelessData3a$avgSigQual, main="Scatterplot Example")

#, 
 #  xlab="Duration Mins", ylab="Average Kbps ", pch=19)

```



Graveyard
```{r}
# https://stackoverflow.com/questions/25264461/find-local-minimum-in-bimodal-distribution-with-r 
# library(ggplot2)
# # k means stuff
#       sig <- data.frame(signal)
#       colnames(sig) <- "X"
#       km <- kmeans(sig, centers=2)
#       sig$clust <- as.factor(km$cluster)
#       ggplot(sig,  aes(x=X)) + 
#   geom_histogram(aes(fill=clust,y=..count../sum(..count..)),
#                      binwidth=0.5, color="grey50")+
#   stat_density(geom="line", color="red")
```


```{r}
par(mar=c(20,5,2,1))
plot(factor(X2016FallSemWireless$WAPID), las=3, main="Number of Connections by Location", ylab="Frequency")
```

